---
title: "Monitoring Stations"
format: html
logo: the-nature-conservancy.png
---



## Introduction

This interactive map displays monitoring stations established for benthic ecological surveys along the coastal waters of Tanzania. Benthic ecosystems form the foundation of marine biodiversity and productivity, supporting critical habitats including coral reefs, seagrass beds, and soft-sediment communities. 

The monitoring stations shown here represent key sampling locations used to assess and track the health and biodiversity of these vital seafloor communities. By visualizing station locations within defined monitoring zones, this tool enables stakeholders to explore spatial patterns in benthic data and supports evidence-based coastal management and conservation efforts.

Use the interactive features to:
- Toggle between different map layers (satellite imagery, street maps)
- Explore monitoring zones and station locations
- Draw custom areas of interest
- Export your annotations as GeoJSON files


```{r}
require(sf)
require(tidyverse)
require(gt)
require(leaflet)
require(leaflet.extras)
require(leafpop)

```


## General

```{r}
cfmas = st_read('cfmas_mainland_pemba.gpkg', quiet = TRUE)
cfmas  = cfmas |> 
    st_make_valid() |> 
    mutate(
        management = if_else(source == 'MCC', "CMG", 'CFMA'),
        area_km2 = as.numeric(st_area(geom)) / 1e6,
        area_ha = as.numeric(st_area(geom)) / 1e4
        )
cmg = cfmas |> 
    filter(management == 'CMG')

cfma = cfmas |> 
    filter(!management == 'CMG')

```


```{r}

# cfma |> 
#     leaflet() |>
#   addProviderTiles(providers$OpenStreetMap, group = "OSM") |>
#   addPolygons()

# you receive this error: ! Don't know how to get polygon data from object of class XY,GEOMETRYCOLLECTION,sfg

# Now correct with st_cast
cfma <- cfma |>
  st_make_valid() |>
  st_cast("MULTIPOLYGON", warn = FALSE) |>
  filter(!st_is_empty(geom))

cfma |> 
    leaflet() |>
  addProviderTiles(providers$OpenStreetMap, group = "OSM") |>
  addPolygons()
```


```{r}
landing_site = st_read('landing_sites.gpkg', quiet = TRUE) |> 
    select(name = Name, village = Village, district = District) |> 
    mutate(
        lon = st_coordinates(geom)[, 1],
        lat = st_coordinates(geom)[, 2]
    )

```

```{r}
mma = st_read('mma.gpkg', quiet = TRUE) |> 
    st_make_valid() |> 
    janitor::clean_names() |> 
    select(name, management = class, region = code) |> 
    mutate(
        centroid = st_centroid(geom),
        lon = st_coordinates(centroid)[, 1],
        lat = st_coordinates(centroid)[, 2],
        area_km2 = as.numeric(st_area(geom)) / 1e6,
        area_ha = as.numeric(st_area(geom)) / 1e4
    )

```


## Allen Coral

```{r}
 benthic_clean =  st_read('allen_benthic_simplified_mapshaper.gpkg', quiet = T)

 # Enhanced color palette
benthic_colors <- c(
  "Coral/Algae" = "#ff6161", 
  "Seagrass" = "#668438",
  "Microalgal Mats" = "#9bcc4f", 
  "Rock" = "#b19c3a",
  "Rubble" = "#e0d05e", 
  "Sand" = "#ffffbe"
)

pal_benth <- colorFactor(palette = benthic_colors, domain = benthic_clean$class) 
benth_classes = benthic_clean |> st_drop_geometry() |> distinct(class) |> pull(class)

```


## Monitoring Stations

The map below provides a comprehensive spatial overview of the established monitoring network. Each station is categorized by its respective management zone, with interactive markers providing real-time insights into ecological indicators such as coral cover percentages and bleaching severity levels. This spatial distribution is critical for identifying vulnerable hotspots and ensuring that conservation resources are allocated effectively across the diverse marine landscapes of Tanzania.


```{r}
stations = read_csv('monitoring_stations.csv') |> 
    janitor::clean_names() |> 
    mutate(
        cover_pct = sample(10:90, n(), replace = TRUE),
        bleaching_level = sample(c("Low", "Moderate", "High", "Severe"), n(), replace = TRUE)
           )
    # |> 
    # mutate(zone = str_remove_all(zone, 'ZONE'))

stations_sf =  stations |> 
    st_as_sf(
        coords = c('lon', 'lat'), 
        crs = 4326, 
        remove = FALSE
        )

# stations_sf |> st_write('monitoring_stations.gpkg', append = FALSE)    

```

```{r}
zones = st_read('monitoring_zones.gpkg') |> 
    mutate(zone = str_replace_all(zone, 'Zone', 'ZONE'))
```


```{r}
pal <- colorFactor(palette = "Set2", domain = stations_sf$zone)

leaflet(stations_sf)  |>
    addProviderTiles(providers$OpenStreetMap, group = "OSM")  |> 
    addProviderTiles(providers$Esri.WorldImagery, group = "Satellite")|> 
    addProviderTiles(providers$CartoDB.Positron, group = "Light") |> 
    addScaleBar(position = "bottomleft") |>
    addFullscreenControl() |>
    addMeasure(
        position = "bottomleft",
        primaryLengthUnit = "kilometers", secondaryLengthUnit = "meters",
        primaryAreaUnit = "sqmeters", secondaryAreaUnit = "hectares",
        activeColor = "#2c3e50",
        completedColor = "#e74c3c",
        localization = "en",
        captureZIndex = 10000
        ) |> 
    addDrawToolbar(
        targetGroup = "draw",
        editOptions = editToolbarOptions(edit = TRUE, remove = TRUE),
        polylineOptions = drawPolylineOptions(),
        polygonOptions = drawPolygonOptions(),
        circleOptions = drawCircleOptions(),
        rectangleOptions = drawRectangleOptions(),
        markerOptions = drawMarkerOptions(),
        circleMarkerOptions = drawCircleMarkerOptions()
        ) |>
    addMiniMap(tiles = providers$CartoDB.Positron, toggleDisplay = TRUE) |> 
    addEasyButton(easyButton(
        icon="fa-crosshairs", title="Locate Me",
        onClick=JS("function(btn, map){ map.locate({setView: true}); }"))) |> 
    addEasyButton(easyButton(
        icon="fa-globe", title="Reset View",
        onClick=JS("function(btn, map){ map.setView([-6.5, 39.3], 8); }")))  |> 
    addPolygons(
        data = zones,
        fillColor = ~pal(zone),
        fillOpacity = 0.2,
        weight = 2,
        color = "black",
        label = ~zone,
        group = "Zones"
    ) |> 
    addCircleMarkers(
        fillColor = ~pal(zone), 
        group = 'Station',
        fillOpacity = 0.8,
        color = "white",
        weight = 2,
        radius = 8,
        label = ~station,
        labelOptions = labelOptions(
            style = list("font-weight" = "bold", padding = "3px 8px"),
            textsize = "15px",
            direction = "auto"
        ),
        popup = ~paste0(
            "<div class='station-popup'>",
            "<div class='popup-header'>", station, "</div>",
            "<div class='popup-content'>",
            "<div class='popup-row'><span class='popup-label'>📍 Location:</span> <span class='popup-value'>", location, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🏷️ MMA:</span> <span class='popup-value'>", mma, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🌍 Zone:</span> <span class='popup-value'>", zone, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📐 Longitude:</span> <span class='popup-value'>", round(lon, 4), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📐 Latitude:</span> <span class='popup-value'>", round(lat, 4), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🌡️ Bleaching:</span> <span class='popup-value' style='color:", 
                case_when(
                    bleaching_level == "Low" ~ "#27ae60",
                    bleaching_level == "Moderate" ~ "#f3d912ff",
                    bleaching_level == "High" ~ "#f8a11fff",
                    bleaching_level == "Severe" ~ "#e74c3c",
                    TRUE ~ "#0c0d0dff"
                ), "; font-weight:bold;'>", bleaching_level, "</span></div>",
            "<div class='popup-row' style='margin-top:2px; margin-bottom:10px;'>",
                "<div style='display:flex; width:100%; height:10px; background:#eee; border-radius:4px; overflow:hidden; border:1px solid #ddd;'>",
                    "<div style='width:25%; background:", ifelse(bleaching_level=="Low", "#27ae60", "#eee"), "; border-right:1px solid white;'></div>",
                    "<div style='width:25%; background:", ifelse(bleaching_level=="Moderate", "#f39c12", "#eee"), "; border-right:1px solid white;'></div>",
                    "<div style='width:25%; background:", ifelse(bleaching_level=="High", "#e67e22", "#eee"), "; border-right:1px solid white;'></div>",
                    "<div style='width:25%; background:", ifelse(bleaching_level=="Severe", "#e74c3c", "#eee"), ";'></div>",
                "</div>",
            "</div>",
            "<div class='popup-row' style='margin-top:10px; flex-direction:column; align-items:flex-start;'>",
                "<span class='popup-label'>🌿 Coral Cover: ", cover_pct, "%</span>",
                "<div style='background:#eee; border-radius:10px; width:100%; height:12px; margin-top:5px;'>",
                    "<div style='background:", 
                        ifelse(cover_pct <= 30, "#e74c3c", 
                        ifelse(cover_pct <= 70, "#f39c12", "#27ae60")), 
                        "; width:", cover_pct, "%; height:100%; border-radius:10px;'>",
                    "</div>",
                "</div>",
            "</div>",
            "</div></div>"
        )
    ) |>
    addLegend(
        pal = pal, 
        values = ~zone, 
        title = "Zone",
        position = "bottomright",
        opacity = 1
    ) |> 
    addLayersControl(
        baseGroups = c("OSM", "Satellite", "Light"),
        overlayGroups = c("Station", "Zones"),
        options = layersControlOptions(collapsed = FALSE)
    )  |> 
    addSearchFeatures(
        targetGroup = "Station",
        options = searchFeaturesOptions(
            zoom = 12, 
            openPopup = TRUE
        )
        ) |> 
    clearSearchFeatures() |>
    htmlwidgets::onRender("
        function(el, x) {
            var map = this;
            var drawnItems = new L.FeatureGroup();
            map.addLayer(drawnItems);
            map.on('draw:created', function (e) {
                drawnItems.addLayer(e.layer);
            });
            L.easyButton('fa-download', function(btn, map){
                var data = drawnItems.toGeoJSON();
                var blob = new Blob([JSON.stringify(data)], {type: 'application/json'});
                var url = URL.createObjectURL(blob);
                var a = document.createElement('a');
                a.href = url;
                a.download = 'drawn_polygons.geojson';
                a.click();
            }, 'Export Drawn Polygons').addTo(map);
        }
    ") |>
    htmlwidgets::onRender("
        function(el, x) {
            var style = document.createElement('style');
            style.textContent = `
                .station-popup {
                    font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
                    background: #ffffff;
                    border-radius: 8px;
                    padding: 15px;
                    box-shadow: 0 4px 12px rgba(0,0,0,0.15);
                    min-width: 250px;
                }
                
                .popup-header {
                    font-size: 30px;
                    font-weight: bold;
                    color: #2c3e50;
                    padding-bottom: 10px;
                    border-bottom: 1px solid #2c3e50;
                    margin-bottom: 10px;
                }
                
                .popup-content {
                    display: flex;
                    flex-direction: column;
                    gap: 8px;
                }
                
                .popup-row {
                    display: flex;
                    justify-content: space-between;
                    align-items: center;
                    font-size: 13px;
                    animation: slideIn 0.3s ease-out;
                }
                
                .popup-label {
                    font-weight: 600;
                    color: #34495e;
                    margin-right: 8px;
                }
                
                .popup-value {
                    color: #0c0d0dff;
                    font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
                }
                
                @keyframes slideIn {
                    from {
                        opacity: 0;
                        transform: translateX(-10px);
                    }
                    to {
                        opacity: 1;
                        transform: translateX(0);
                    }
                }
                
                .leaflet-popup-content {
                    margin: 0 !important;
                    padding: 0 !important;
                }
                
                .leaflet-popup {
                    animation: popupSlideUp 0.3s cubic-bezier(0.34, 1.56, 0.64, 1);
                }
                
                @keyframes popupSlideUp {
                    from {
                        opacity: 0;
                        transform: translateY(20px);
                    }
                    to {
                        opacity: 1;
                        transform: translateY(0);
                    }
                }
                
                .leaflet-label {
                    background: rgba(44, 62, 80, 0.9);
                    color: white;
                    border-radius: 4px;
                    box-shadow: 0 2px 8px rgba(0,0,0,0.2);
                    font-weight: 600;
                    animation: labelFadeIn 0.2s ease-out;
                }
                
                @keyframes labelFadeIn {
                    from {
                        opacity: 0;
                        transform: scale(0.9);
                    }
                    to {
                        opacity: 1;
                        transform: scale(1);
                    }
                }
                
                .leaflet-marker-icon:hover {
                    filter: drop-shadow(0 0 8px rgba(52, 152, 219, 0.6));
                    transform: scale(1.1);
                    transition: all 0.2s ease;
                }
            `;
            document.head.appendChild(style);
            
            // Add hover effects to circle markers
            var map = this;
            var originalRadius = 8;
            var expandedRadius = 12;

            // Welcome Popup Guide
            var guideContent = `
                <div style='font-family: sans-serif; padding: 10px; min-width: 200px;'>
                    <h1 style='margin-top:0; color: #2c3e50;'>🗺️ Benthic Inventory Dashboard</h1>
                    <p>Explore integrated ecological data across Tanzania's coastal waters:</p>
                    <ul style='max-height: 150px; overflow-y: auto; padding-right: 5px;'>
                        <li><b>Integrated Data:</b> Toggle between <b>Long-term Monitoring Stations</b> and high-resolution <b>BES Grid Points</b> using the layer control.</li>
                        <li><b>Spatial Navigation:</b> Pan across the coastline and zoom into specific Marine Managed Areas (MMAs) or monitoring zones.</li>
                        <li><b>Station Search:</b> Use the 🔍 icon to quickly locate specific monitoring sites by name.</li>
                        <li><b>Ecological Insights:</b> Click any marker to view detailed metrics including coral cover %, bleaching severity, and substrate composition.</li>
                        <li><b>Environmental Context:</b> Switch between Satellite imagery and Street maps to analyze habitat surroundings.</li>
                        <li><b>Benthic Habitats:</b> These layers are derived from Allen Coral Atlas data, simplified for rapid web performance. Please use these specific layers with caution for high-precision spatial analysis.</li>
                        <li><b>Custom Analysis:</b> Use the drawing toolbar to delineate areas of interest and export coordinates as GeoJSON for external GIS analysis.</li>
                    </ul>
                    <hr><small>Developed by: <b>Masumbuko Semba</b> creditied to: Gemini, Claude, and Raptor AI via GitHub Copilot using Positron IDE in R language.</small>
                </div>`;
            
            var guidePopup = L.popup({
                autoPan: false,
                className: 'floating-guide'
            })
            .setLatLng(map.getCenter())
            .setContent(guideContent);
            
            L.easyButton('fa-info-circle', function(btn, map){
                guidePopup.openOn(map);
            }, 'Show Dashboard Guide').addTo(map);

            guidePopup
                .openOn(map);

            // CSS to make it float and stay fixed during pans
            var floatStyle = document.createElement('style');
            floatStyle.textContent = `
                .floating-guide {
                    position: fixed !important;
                    top: 100px !important;
                    left: 50px !important;
                    transform: none !important;
                    z-index: 1000;
                }
            `;
            document.head.appendChild(floatStyle);

            if (map._popup && map._popup._container) {
                var container = map._popup._container;
                var draggable = new L.Draggable(container);
                draggable.enable();
                container.style.cursor = 'move';
                draggable.on('drag', function(e) {
                    container.style.transform = e.target._newPos;
                });
            }

            map.on('draw:created', function (e) {
                drawnItems.addLayer(e.layer);
            });
        }
    ") 

```


```{r}
stations |> 
    group_by(location, zone) |> 
    summarize( n = n()) |> 
    ungroup() |>
    arrange(zone) |> 
    add_row(location = 'BES', zone = "OPEN", n = 30) |> 
    gt::gt()
```



```{r}
stations |> 
    group_by(location, zone) |> 
    summarize(sites = paste(station, collapse = ', ')) |> 
    ungroup() |> 
    left_join(
        stations |> 
            group_by(location, zone) |> 
            summarize(Total = n()) |> 
            ungroup()
    ) |> 
    arrange(zone) |> 
    add_row(location = 'BES', zone = "OPEN", Total = 30) |> 
    gt::gt() |> 
    gt::cols_width(
        sites ~ gt::pct(55),
        everything()~pct(15)
    ) |> 
    tab_options(
        latex.use_longtable = TRUE, 
        data_row.padding = px(5), 
        table.font.size = '11pt', 
        table.width = pct(100)
    )

```


## Benthic Ecological Surveys (BES)

The Benthic Ecological Survey (BES) stations represent a specialized subset of the monitoring network, specifically designed to capture high-resolution data on seafloor composition and health. These surveys utilize standardized transect methodologies to quantify the cover of live coral, macroalgae, and other substrate types. By integrating these detailed biological assessments with the broader spatial data from the monitoring zones, the BES framework provides the empirical evidence necessary to track long-term shifts in community structure and evaluate the resilience of Tanzania's coastal ecosystems against climate-induced stressors.



## Labelset

```{r}
labelset = read_csv('labelset.csv')

labelset = labelset |>
  mutate(
    ecological_group = case_when(
      `Functional Group` %in% c("Hard coral") ~ "Live Coral",
      `Functional Group` %in% c("Seagrass") ~ "Seagrass",
      `Functional Group` %in% c("Algae", "Other Invertebrates") ~ "Other Biota",
      `Functional Group` %in% c("Soft Substrate", "Hard Substrate", "Other") ~ "Substrate"
    )
  ) 

```

```{r}
#| label: tbl-labelset2
#| tbl-cap: Functional and ecological labelset  used to characterize the benthic community and other substralte


 labelset  |>
#   select(Name, `Functional Group`, ecological_group) |>
  gt() |>
  opt_row_striping() |>
  tab_options(
    latex.use_longtable = TRUE,
    data_row.padding = px(5),
    table.font.size = '10pt'
  ) |> 
    cols_width(
        Name ~ '3in'
    )
```




```{r}
bes = read_csv('BES_site_full.csv') |> 
rename(seascape = Region) |> 
mutate(
    seascape = str_replace_all(seascape, 'Mainland_South', 'Mainland'),
    seascape = str_replace_all(seascape, 'Mainland_North', 'Mainland'),
    seascape = if_else(seascape == 'Mainland', 'Mainland', 'Zanzibar')
)

```



```{r}
bes_long =  bes |> 
    select(-1,-2) |> 
    pivot_longer(
        cols = 5:21, 
        names_to = "Short Code", 
        values_to = 'pct'
    ) |>
    left_join(labelset, by = 'Short Code') |>
    # Extract the numeric part of the image filename
    mutate(grid = as.integer(str_extract(Image, "\\d+"))) |>
    # # The filter below is to check for images where the number was not extracted
    # filter(is.na(grid)) |> 
    separate(Image, c('Image', 'bb'), sep = -4)|> 
    separate(Image, c('Image', 'replicate'), sep = -1) |> 
    mutate(
        replicate = str_to_upper(replicate),
        pct = na_if(pct, 0)
        )
        
bes_long_clean = bes_long |>     
    select(-Image, -bb, -`Functional Group`, -`Short Code`) |> 
    relocate(grid, .after = seascape) |> 
    relocate(pct, .after = ecological_group) |> 
    filter(!is.na(pct))|> 
janitor::clean_names()


```


```{r} 
set.seed(123)

bes_coral_seagrass = bes_long_clean |> 
  filter(ecological_group %in% c('Live Coral', 'Seagrass')) |>
  filter(!is.na(longitude )) |> 
  select(-name)


bes_coral_seagrass_sf = bes_coral_seagrass  |> 
    filter(pct > 95) |> 
    group_by(seascape, ecological_group) |> 
    sample_n(12) |> 
    ungroup() |> 
    # distinct(ecological_group)
  st_as_sf(coords = c('longitude', 'latitude'), crs = 4326, remove = FALSE) 
  
bes_coral_seagrass_sf|> 
  mapview::mapview()


```


## Revised map (Monitoring + BES)


```{r}
pal <- colorFactor(palette = "Set2", domain = stations_sf$zone)
pal2 <- colorFactor(palette = "Set1", domain = bes_coral_seagrass_sf$ecological_group)


map = leaflet() |> 
    # login window
    htmlwidgets::onRender("
        function(el, x) {
            var map = this;
            var loginHTML = \`
                <div style=\"
                    position: fixed;
                    top: 0;
                    left: 0;
                    width: 100%;
                    height: 100%;
                    background: rgba(0, 0, 0, 0.2);
                    display: flex;
                    align-items: center;
                    justify-content: center;
                    z-index: 10000;
                    font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
                \" id=\"loginOverlay\">
                    <div style=\"
                        background: #ffffff;
                        border-radius: 12px;
                        padding: 0;
                        width: 100%;
                        max-width: 450px;
                        box-shadow: 0 20px 40px rgba(0,0,0,0.2);
                        overflow: hidden;
                        animation: slideUp 0.4s cubic-bezier(0.34, 1.56, 0.64, 1);
                    \">
                        <div style=\"
                            background: #1a2a3a;
                            padding: 40px 30px 30px;
                            text-align: center;
                            color: white;
                            border-bottom: 4px solid #2a7fffff;
                        \">
                            <div style=\"
                                font-size: 68px;
                                margin-bottom: 15px;
                            \">🗺️</div>
                            <h2 style=\"
                                margin: 0;
                                font-size: 28px;
                                font-weight: 700;
                                letter-spacing: -0.5px;
                                color: #80b3ffff;
                            \">Benthic Inventory</h2>
                            <p style=\"
                                margin: 8px 0 0 0;
                                font-size: 14px;
                                color: #d5e5ffff;
                                font-weight: 300;
                            \">Tanzania Benthic Ecosystem Dashboard</p>
                        </div>
                        
                        <div style=\"
                            background: #ffffff;
                            padding: 40px 30px;
                        \">
                            <div style=\"margin-bottom: 20px;\">
                                <label style=\"
                                    display: block;
                                    font-size: 12px;
                                    font-weight: 600;
                                    color: #1a2a3a;
                                    margin-bottom: 8px;
                                    text-transform: uppercase;
                                    letter-spacing: 1px;
                                \">Username</label>
                                <input type=\"text\" id=\"loginUser\" placeholder=\"Enter your username\" style=\"
                                    width: 100%;
                                    padding: 12px 15px;
                                    border: 2px solid #1a2a3a;
                                    border-radius: 4px;
                                    font-size: 14px;
                                    transition: all 0.3s ease;
                                    box-sizing: border-box;
                                \" onfocus=\"this.style.borderColor='#00d1b2';\" />
                            </div>
                            
                            <div style=\"margin-bottom: 30px;\">
                                <label style=\"
                                    display: block;
                                    font-size: 12px;
                                    font-weight: 600;
                                    color: #1a2a3a;
                                    margin-bottom: 8px;
                                    text-transform: uppercase;
                                    letter-spacing: 1px;
                                \">Password</label>
                                <input type=\"password\" id=\"loginPass\" placeholder=\"Enter your password\" style=\"
                                    width: 100%;
                                    padding: 12px 15px;
                                    border: 2px solid #1a2a3a;
                                    border-radius: 4px;
                                    font-size: 14px;
                                    transition: all 0.3s ease;
                                    box-sizing: border-box;
                                \" onfocus=\"this.style.borderColor='#00d1b2';\" />
                            </div>
                            
                            <div style=\"
                                display: grid;
                                grid-template-columns: 1fr 1fr;
                                gap: 12px;
                            \">
                                <button id=\"loginBtn\" style=\"
                                    background: #5599ffff;
                                    color: white;
                                    border: none;
                                    padding: 12px 24px;
                                    border-radius: 4px;
                                    font-size: 14px;
                                    font-weight: 700;
                                    cursor: pointer;
                                    transition: all 0.3s ease;
                                    text-transform: uppercase;
                                    letter-spacing: 0.5px;
                                \" onmouseover=\"this.style.background='#d5e5ffff'\" onmouseout=\"this.style.background='#5599ffff'\">Sign In</button>
                                
                            </div>
                            
                            <div style=\"
                                margin-top: 20px;
                                padding-top: 20px;
                                border-top: 2px solid #f0f0f0;
                                font-size: 12px;
                                color: #1a2a3a;
                                line-height: 1.6;
                            \">

                                    <p style=\"margin: 0 0 8px 0;\"><strong>🔐 Accessing the Tool:</strong></p>
                                    <p style=\"margin: 5px 0; font-size: 11px; color: #666;\">Please contact the developer to request access credentials.</p>
                                </div>

                            </div>
                        </div>
                    </div>
                </div>
            \`;
            document.body.insertAdjacentHTML('beforeend', loginHTML);
            
 // Usability Tracking Function
var trackEvent = function(action, user) {
    var timestamp = new Date().toISOString();
    var logEntry = { event: action, username: user, time: timestamp, screen: window.innerWidth + 'x' + window.innerHeight };
    
    // Local logging for debugging
    localStorage.setItem('last_session_event', JSON.stringify(logEntry));
    console.log('Event tracked:', logEntry);

    // Send to Google Form - CORRECTED URL
    var formId = '1TT1BhEwo9IlbOvL-mKkTrKGuEYTz--t1y-_SAzopEVQ';
    var formUrl = 'https://docs.google.com/forms/d/e/' + formId + '/formResponse';
    
    var formData = new FormData();
    // You MUST update these entry IDs with your actual form field IDs
    formData.append('entry.hj99tb2', user);      // Replace with your 'user' field ID
    formData.append('entry.hj99tb4', action);    // Replace with your 'action' field ID
    formData.append('entry.hj99tb6', timestamp); // Replace with your 'timestamp' field ID
    
    fetch(formUrl, { 
        method: 'POST', 
        mode: 'no-cors', 
        body: formData 
    }).catch(err => console.log('Form submission sent'));
};

            // Add event listeners for login buttons
            var loginBtn = document.getElementById('loginBtn');
            var loginOverlay = document.getElementById('loginOverlay');
            
            loginBtn.addEventListener('click', function() {
                var user = document.getElementById('loginUser').value;
                var pass = document.getElementById('loginPass').value;
                
                var validUsers = {
                    'semba': 'semba',
                    'asuta': 'asuta',
                    'tnc': 'tnc',
                    'user': 'guest'
                };
                
                if (validUsers[user] && validUsers[user] === pass) {
                    loginOverlay.style.display = 'none';
                    trackEvent('login_success', user);
                } else if (!user || !pass) {
                    alert('Please enter both username and password');
                } else {
                    alert('Invalid credentials. Please contact Semba for access.');
                }
            });

            // Track when user leaves the dashboard (Logout/Close)
            window.addEventListener('beforeunload', function() {
                var user = document.getElementById('loginUser').value || 'unknown';
                trackEvent('session_end', user);
            });
        }
    ") |>
        # Basemap header insert
    htmlwidgets::onRender("
        function(el, x) {
            var layerControl = document.querySelector('.leaflet-control-layers-base');
            if (layerControl) {
                var header = document.createElement('div');
                header.style.fontWeight = 'bold';
                header.style.color = '#2c3e50';
                header.style.margin = '0 0 5px 0';
                header.style.fontSize = '12px';
                header.innerHTML = '🗺️ BASEMAPS';
                layerControl.parentNode.insertBefore(header, layerControl);
            }
        }") |>
    addProviderTiles(providers$OpenStreetMap, group = "<span style='color: #2c3e50; font-weight: bold;'>🗺️ Street Map</span>")  |> 
    addProviderTiles(providers$Esri.WorldImagery, group = "<span style='color: #2c3e50; font-weight: bold;'>🛰️ Satellite</span>")|> 
    addProviderTiles(providers$CartoDB.Positron, group = "<span style='color: #2c3e50; font-weight: bold;'>⚪ Minimal Light</span>") |> 
    addScaleBar(position = "bottomleft") |>
    addFullscreenControl() |>
    addMeasure(
        position = "bottomleft",
        primaryLengthUnit = "kilometers", secondaryLengthUnit = "meters",
        primaryAreaUnit = "sqmeters", secondaryAreaUnit = "hectares",
        activeColor = "#2c3e50",
        completedColor = "#e74c3c",
        localization = "en",
        captureZIndex = 10000
        ) |> 
    addDrawToolbar(
        targetGroup = "draw",
        editOptions = editToolbarOptions(edit = TRUE, remove = TRUE),
        polylineOptions = drawPolylineOptions(),
        polygonOptions = drawPolygonOptions(),
        circleOptions = drawCircleOptions(),
        rectangleOptions = drawRectangleOptions(),
        markerOptions = drawMarkerOptions(),
        circleMarkerOptions = drawCircleMarkerOptions()
        ) |>
    addMiniMap(tiles = providers$CartoDB.Positron, toggleDisplay = TRUE, position = "bottomleft") |> 
    addEasyButton(easyButton(
        icon="fa-crosshairs", title="Locate Me",
        onClick=JS("function(btn, map){ map.locate({setView: true}); }"))) |> 
    addEasyButton(easyButton(
        icon="fa-globe", title="Reset View",
        onClick=JS("function(btn, map){ map.setView([-6.5, 39.3], 8); }")))  |> 
    addPolygons(
        data = zones,
        fillColor = ~pal(zone),
        fillOpacity = 0.2,
        weight = 2,
        color = "black",
        label = ~zone,
        group = "Zones"
    ) |> 
    addPolygons(
        data = cmg,
        fillColor = "#27ae60",
        fillOpacity = 0.1,
        weight = 1,
        color = "#27ae60",
        dashArray = "3",
        label = ~name,
        popup = ~paste0(
            "<div class='station-popup'>",
            "<div class='popup-header'>", name, "</div>",
            "<div class='popup-content'>",
            "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", management, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", round(area_km2, 2), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", round(area_ha, 2), "</span></div>",
            "</div></div>"
        ),
        group = "CMG"
    ) |> 
    addPolygons(
        data = cfma,
        fillColor = "#3498db",
        fillOpacity = 0.1,
        weight = 1,
        color = "#3498db",
        dashArray = "3",
        label = ~name,
        popup = ~paste0(
            "<div class='station-popup'>",
            "<div class='popup-header'>", name, "</div>",
            "<div class='popup-content'>",
            "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", management, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", round(area_km2, 2), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", round(area_ha, 2), "</span></div>",
            "</div></div>"
        ),
        group = "CFMA"
    ) |> 
    addPolygons(
        data = mma,
        fillColor = "#9b59b6",
        fillOpacity = 0.1,
        weight = 1,
        color = "#9b59b6",
        dashArray = "3",
        label = ~name,
        popup = ~paste0(
            "<div class='station-popup'>",
            "<div class='popup-header'>", name, "</div>",
            "<div class='popup-content'>",
            "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", management, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>Region:</span> <span class='popup-value'>", region, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", round(area_km2, 2), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", round(area_ha, 2), "</span></div>",
            "</div></div>"
        ),
        group = "MMA"
    ) |> 
    addCircleMarkers( 
        data = bes_coral_seagrass_sf,
        fillColor = ~pal2(ecological_group), 
        group = 'BES',
        fillOpacity = 0.8,
        color = "white",
        weight = 2,
        radius = 8,
        label = ~paste("Grid:", grid),
        labelOptions = labelOptions(
            style = list("font-weight" = "bold", padding = "3px 8px"),
            textsize = "15px",
            direction = "auto"
        ),
        popup = ~paste0(
            "<div class='station-popup'>",
            "<div class='popup-header'>BES Grid: ", grid, "</div>",
            "<div class='popup-content'>",
            # "<div class='popup-row'><span class='popup-label'>🌊 Seascape:</span> <span class='popup-value'>", seascape, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🌿 Group:</span> <span class='popup-value'>", ecological_group, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📊 Coverage:</span> <span class='popup-value'>", round(pct, 1), "%</span></div>",
            "<div class='popup-row'><span class='popup-label'>🔄 Replicate:</span> <span class='popup-value'>", replicate, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📐 Longitude:</span> <span class='popup-value'>", round(longitude, 4), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📐 Latitude:</span> <span class='popup-value'>", round(latitude, 4), "</span></div>",
            "<div class='popup-row' style='margin-top:10px; flex-direction:column; align-items:flex-start;'>",
                "<div style='background:#eee; border-radius:10px; width:100%; height:12px; margin-top:5px;'>",
                    "<div style='background:#3498db; width:", pct, "%; height:100%; border-radius:10px;'></div>",
                "</div>",
            "</div>",
            "</div></div>"
        )
        ) |> 
    addCircleMarkers(
        data = landing_site,
        group = "Landing Sites",
        fillColor = "#000000ff",
        fillOpacity = 0.7,
        color = "white",
        weight = 1,
        radius = 5,
        label = ~name,
        popup = ~paste0(
            "<div class='station-popup'>",
            "<div class='popup-header'>Landing Site</div>",
            "<div class='popup-content'>",
            "<div class='popup-row'><span class='popup-label'>⚓ Name:</span> <span class='popup-value'>", name, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🏘️ Village:</span> <span class='popup-value'>", village, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🏛️ District:</span> <span class='popup-value'>", district, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📐 Longitude:</span> <span class='popup-value'>", round(lon, 4), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📐 Latitude:</span> <span class='popup-value'>", round(lat, 4), "</span></div>",
            "</div></div>"
        )
    ) |>
    addCircleMarkers( 
        data = stations_sf,
        fillColor = ~pal(zone), 
        group = 'Station',
        fillOpacity = 0.8,
        color = "white",
        weight = 2,
        radius = 8,
        label = ~station,
        labelOptions = labelOptions(
            style = list("font-weight" = "bold", padding = "3px 8px"),
            textsize = "15px",
            direction = "auto"
        ),
        popup = ~paste0(
            "<div class='station-popup'>",
            "<div class='popup-header'>", station, "</div>",
            "<div class='popup-content'>",
            "<div class='popup-row'><span class='popup-label'>📍 Location:</span> <span class='popup-value'>", location, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🏷️ MMA:</span> <span class='popup-value'>", mma, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🌍 Zone:</span> <span class='popup-value'>", zone, "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📐 Longitude:</span> <span class='popup-value'>", round(lon, 4), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>📐 Latitude:</span> <span class='popup-value'>", round(lat, 4), "</span></div>",
            "<div class='popup-row'><span class='popup-label'>🌡️ Bleaching:</span> <span class='popup-value' style='color:", 
                case_when(
                    bleaching_level == "Low" ~ "#27ae60",
                    bleaching_level == "Moderate" ~ "#f3d912ff",
                    bleaching_level == "High" ~ "#f8a11fff",
                    bleaching_level == "Severe" ~ "#e74c3c",
                    TRUE ~ "#0c0d0dff"
                ), "; font-weight:bold;'>", bleaching_level, "</span></div>",
            "<div class='popup-row' style='margin-top:2px; margin-bottom:10px;'>",
                "<div style='display:flex; width:100%; height:10px; background:#eee; border-radius:4px; overflow:hidden; border:1px solid #ddd;'>",
                    "<div style='width:25%; background:", ifelse(bleaching_level=="Low", "#27ae60", "#eee"), "; border-right:1px solid white;'></div>",
                    "<div style='width:25%; background:", ifelse(bleaching_level=="Moderate", "#f39c12", "#eee"), "; border-right:1px solid white;'></div>",
                    "<div style='width:25%; background:", ifelse(bleaching_level=="High", "#e67e22", "#eee"), "; border-right:1px solid white;'></div>",
                    "<div style='width:25%; background:", ifelse(bleaching_level=="Severe", "#e74c3c", "#eee"), ";'></div>",
                "</div>",
            "</div>",
            "<div class='popup-row' style='margin-top:10px; flex-direction:column; align-items:flex-start;'>",
                "<span class='popup-label'>🌿 Coral Cover: ", cover_pct, "%</span>",
                "<div style='background:#eee; border-radius:10px; width:100%; height:12px; margin-top:5px;'>",
                    "<div style='background:", 
                        ifelse(cover_pct <= 30, "#e74c3c", 
                        ifelse(cover_pct <= 70, "#f39c12", "#27ae60")), 
                        "; width:", cover_pct, "%; height:100%; border-radius:10px;'>",
                    "</div>",
                "</div>",
            "</div>",
            "</div></div>"
        )
    )  |> # Benthic habitat layers with enhanced popups
  addPolygons(
    data = benthic_clean |> filter(class == benth_classes[1]),
    group = benth_classes[1], 
    fillColor = unname(benthic_colors[as.character(benth_classes[1])]), 
    stroke = TRUE,
    color = "#ffffff",
    weight = 0.5,
    opacity = 0.3,
    fillOpacity = 1.0,
    smoothFactor = 1,
    popup = ~paste0(
        "<div class='station-popup'>",
        "<div class='popup-header'>", benth_classes[1], "</div>",
         "<div class='popup-content'>",
        "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", 'Benthic Habitat', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Region:</span> <span class='popup-value'>", 'Tanzania', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "</div></div>"
    ),
    highlightOptions = highlightOptions(weight = 2, fillOpacity = 0.9)
  ) |> 
  addPolygons(
    data = benthic_clean |> filter(class == benth_classes[2]),
    group = benth_classes[2], 
    fillColor = unname(benthic_colors[as.character(benth_classes[2])]), 
    stroke = TRUE,
    color = "#ffffff",
    weight = 0.5,
    opacity = 0.3,
    fillOpacity = 1.0,
    smoothFactor = 1,
    popup = ~paste0(
        "<div class='station-popup'>",
        "<div class='popup-header'>", benth_classes[2], "</div>",
         "<div class='popup-content'>",
        "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", 'Benthic Habitat', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Region:</span> <span class='popup-value'>", 'Tanzania', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "</div></div>"
    ),
    highlightOptions = highlightOptions(weight = 2, fillOpacity = 0.9)
  ) |> 
  addPolygons(
    data = benthic_clean |> filter(class == benth_classes[3]),
    group = benth_classes[3], 
    fillColor = unname(benthic_colors[as.character(benth_classes[3])]), 
    stroke = TRUE,
    color = "#ffffff",
    weight = 0.5,
    opacity = 0.3,
    fillOpacity = 1.0,
    smoothFactor = 1,
    popup = ~paste0(
        "<div class='station-popup'>",
        "<div class='popup-header'>", benth_classes[3], "</div>",
         "<div class='popup-content'>",
        "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", 'Benthic Habitat', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Region:</span> <span class='popup-value'>", 'Tanzania', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "</div></div>"
    ),
    highlightOptions = highlightOptions(weight = 2, fillOpacity = 0.9)
  ) |> 
  addPolygons(
    data = benthic_clean |> filter(class == benth_classes[4]),
    group = benth_classes[4], 
    fillColor = unname(benthic_colors[as.character(benth_classes[4])]), 
    stroke = TRUE,
    color = "#ffffff",
    weight = 0.5,
    opacity = 0.3,
    fillOpacity = 1.0,
    smoothFactor = 1,
    popup = ~paste0(
        "<div class='station-popup'>",
        "<div class='popup-header'>", benth_classes[4], "</div>",
         "<div class='popup-content'>",
        "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", 'Benthic Habitat', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Region:</span> <span class='popup-value'>", 'Tanzania', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "</div></div>"
    ),
    highlightOptions = highlightOptions(weight = 2, fillOpacity = 0.9)
  ) |> 
  addPolygons(
    data = benthic_clean |> filter(class == benth_classes[5]),
    group = benth_classes[5], 
    fillColor = unname(benthic_colors[as.character(benth_classes[5])]), 
    stroke = TRUE,
    color = "#ffffff",
    weight = 0.5,
    opacity = 0.3,
    fillOpacity = 1.0,
    smoothFactor = 1,
    popup = ~paste0(
        "<div class='station-popup'>",
        "<div class='popup-header'>", benth_classes[5], "</div>",
         "<div class='popup-content'>",
        "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", 'Benthic Habitat', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Region:</span> <span class='popup-value'>", 'Tanzania', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "</div></div>"
    ),
    highlightOptions = highlightOptions(weight = 2, fillOpacity = 0.9)
  ) |> 
  addPolygons(
    data = benthic_clean |> filter(class == benth_classes[6]),
    group = benth_classes[6], 
    fillColor = unname(benthic_colors[as.character(benth_classes[6])]), 
    stroke = TRUE,
    color = "#ffffff",
    weight = 0.5,
    opacity = 0.3,
    fillOpacity = 1.0,
    smoothFactor = 1,
    popup = ~paste0(
        "<div class='station-popup'>",
        "<div class='popup-header'>", benth_classes[6], "</div>",
         "<div class='popup-content'>",
        "<div class='popup-row'><span class='popup-label'>Type:</span> <span class='popup-value'>", 'Benthic Habitat', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Region:</span> <span class='popup-value'>", 'Tanzania', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (km²):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "<div class='popup-row'><span class='popup-label'>Area (ha):</span> <span class='popup-value'>", 'TBD', "</span></div>",
        "</div></div>"
    ),
    highlightOptions = highlightOptions(weight = 2, fillOpacity = 0.9)
  ) |>
    addLegend(
        data = stations_sf,
        pal = pal, 
        values = ~zone, 
        title = "<strong style='font-size:14px;'>Coral Reef</strong>", 
        position = "bottomright",
        opacity = 1,
        group = 'Zones'
    ) |> 
    addLegend(
        data = bes_coral_seagrass_sf,
        pal = pal2, 
        values = ~ecological_group, 
        title = "<strong style='font-size:14px;'>BES Benthic</strong>", 
        position = "bottomright",
        opacity = 1,
        group = "BES"
    )  |>
  
  # Enhanced legend for benthic habitats
  addLegend(
    pal = pal_benth, 
    values = benthic_clean$class, 
    title = "<strong style='font-size:14px;'>Benthic Habitats</strong>", 
    position = "bottomright", 
    opacity = 0.95,
    labFormat = labelFormat(prefix = ""),
    className = "info legend benthic-legend"
  )  |> 
    addLayersControl(
        baseGroups = c(
            "<span style='color: #2c3e50; font-weight: bold;'>🗺️ Street Map</span>", 
            "<span style='color: #2c3e50; font-weight: bold;'>🛰️ Satellite</span>", 
            "<span style='color: #2c3e50; font-weight: bold;'>⚪ Minimal Light</span>"
        ),
        overlayGroups = c("Zones", "Station",'BES', "Landing Sites",  "CMG", "CFMA", "MMA",
      benth_classes[1], benth_classes[2], benth_classes[3], 
      benth_classes[4], benth_classes[5], benth_classes[6]),
        options = layersControlOptions(collapsed = FALSE)
    )  |> 
    hideGroup(group = c( "Landing Sites",  "CMG", "CFMA", "MMA",
      benth_classes[1], benth_classes[2], benth_classes[3], 
      benth_classes[4], benth_classes[5], benth_classes[6])) |> 
    addSearchFeatures(
        targetGroup = "Station",
        options = searchFeaturesOptions(
            zoom = 12, 
            openPopup = TRUE
        )
        ) |> 
    clearSearchFeatures() |>
    htmlwidgets::onRender("
        function(el, x) {
            var map = this;
            var drawnItems = new L.FeatureGroup();
            map.addLayer(drawnItems);
            map.on('draw:created', function (e) {
                drawnItems.addLayer(e.layer);
            });
            L.easyButton('fa-download', function(btn, map){
                var data = drawnItems.toGeoJSON();
                var blob = new Blob([JSON.stringify(data)], {type: 'application/json'});
                var url = URL.createObjectURL(blob);
                var a = document.createElement('a');
                a.href = url;
                a.download = 'drawn_polygons.geojson';
                a.click();
            }, 'Export Drawn Polygons').addTo(map);
        }
    ") |>
    htmlwidgets::onRender("
        function(el, x) {
            var style = document.createElement('style');
            style.textContent = `
                .station-popup {
                    font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
                    background: #ffffff;
                    border-radius: 8px;
                    padding: 15px;
                    box-shadow: 0 4px 12px rgba(0,0,0,0.15);
                    min-width: 250px;
                }
                
                .popup-header {
                    font-size: 30px;
                    font-weight: bold;
                    color: #2c3e50;
                    padding-bottom: 10px;
                    border-bottom: 1px solid #2c3e50;
                    margin-bottom: 10px;
                }
                
                .popup-content {
                    display: flex;
                    flex-direction: column;
                    gap: 8px;
                }
                
                .popup-row {
                    display: flex;
                    justify-content: space-between;
                    align-items: center;
                    font-size: 13px;
                    animation: slideIn 0.3s ease-out;
                }
                
                .popup-label {
                    font-weight: 600;
                    color: #34495e;
                    margin-right: 8px;
                }
                
                .popup-value {
                    color: #0c0d0dff;
                    font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
                }
                
                @keyframes slideIn {
                    from {
                        opacity: 0;
                        transform: translateX(-10px);
                    }
                    to {
                        opacity: 1;
                        transform: translateX(0);
                    }
                }
                
                .leaflet-popup-content {
                    margin: 0 !important;
                    padding: 0 !important;
                }
                
                .leaflet-popup {
                    animation: popupSlideUp 0.3s cubic-bezier(0.34, 1.56, 0.64, 1);
                }
                
                @keyframes popupSlideUp {
                    from {
                        opacity: 0;
                        transform: translateY(20px);
                    }
                    to {
                        opacity: 1;
                        transform: translateY(0);
                    }
                }
                
                .leaflet-label {
                    background: rgba(44, 62, 80, 0.9);
                    color: white;
                    border-radius: 4px;
                    box-shadow: 0 2px 8px rgba(0,0,0,0.2);
                    font-weight: 600;
                    animation: labelFadeIn 0.2s ease-out;
                }
                
                @keyframes labelFadeIn {
                    from {
                        opacity: 0;
                        transform: scale(0.9);
                    }
                    to {
                        opacity: 1;
                        transform: scale(1);
                    }
                }
                
                .leaflet-marker-icon:hover {
                    filter: drop-shadow(0 0 8px rgba(52, 152, 219, 0.6));
                    transform: scale(1.1);
                    transition: all 0.2s ease;
                }
            `;
            document.head.appendChild(style);
            
            // Add hover effects to circle markers
            var map = this;
            var originalRadius = 8;
            var expandedRadius = 12;

            // Welcome Popup Guide
            var guideContent = `
                <div style='font-family: sans-serif; padding: 10px; min-width: 200px;'>
                    <h1 style='margin-top:0; color: #2c3e50;'>🗺️ Benthic Inventory Dashboard Guide</h1>
                    <p>Explore integrated ecological data across Tanzania's coastal waters:</p>
                    <ul style='max-height: 150px; overflow-y: auto; padding-right: 5px;'>
                        <li><b>Integrated Data:</b> Toggle between <b>Long-term Monitoring Stations</b> and low-resolution <b>BES Grid Points</b> using the layer control.</li>
                        <li><b>Spatial Navigation:</b> Pan across the coastline and zoom into specific Marine Managed Areas (MMAs) or monitoring zones.</li>
                        <li><b>Station Search:</b> Use the 🔍 icon to quickly locate specific monitoring sites by name.</li>
                        <li><b>Ecological Insights:</b> Click any marker to view detailed metrics including coral cover %, bleaching severity, and substrate composition.</li>
                        <li><b>Environmental Context:</b> Switch between Satellite imagery and Street maps to analyze habitat surroundings.</li>
                        <li><b>Custom Analysis:</b> Use the drawing toolbar to delineate areas of interest and export coordinates as GeoJSON for external GIS analysis.</li>
                    </ul>
                    <p style='font-size: 11px; background: #fff3cd; padding: 8px; border-radius: 4px; border-left: 8px solid #ffc107;'><b>Note:</b> Benthic Habitat layers are derived from Allen Coral Atlas data, simplified for rapid web performance. Please use these specific layers with caution for high-precision spatial analysis.</p>
                    <hr>
                    <p style='font-size: 9px; background: #f6ffedff; color: #2c3e50; padding: 8px; border-radius: 4px; border-left: 8px solid #aad400ff;'>Developed by: <b>Masumbuko Semba</b>. Built using <b>R language</b> (sf, tidyverse, leaflet, gt, htmlwidgets), JavaScript, SQL, HTML, and CSS within the Positron IDE and Quarto.</p>
                </div>`;
            
            var guidePopup = L.popup({
                autoPan: false,
                className: 'floating-guide'
            })
            .setLatLng(map.getCenter())
            .setContent(guideContent);
            
            L.easyButton('fa-info-circle', function(btn, map){
                guidePopup.openOn(map);
            }, 'Show Dashboard Guide').addTo(map);

            guidePopup
                .openOn(map);

            // CSS to make it float and stay fixed during pans
            var floatStyle = document.createElement('style');
            floatStyle.textContent = `
                .floating-guide {
                    position: fixed !important;
                    top: 100px !important;
                    left: 50px !important;
                    transform: none !important;
                    z-index: 1000;
                }
            `;
            document.head.appendChild(floatStyle);

            if (map._popup && map._popup._container) {
                var container = map._popup._container;
                var draggable = new L.Draggable(container);
                draggable.enable();
                container.style.cursor = 'move';
                draggable.on('drag', function(e) {
                    container.style.transform = e.target._newPos;
                });
            }

            map.on('draw:created', function (e) {
                drawnItems.addLayer(e.layer);
            });
        }
    ") |>
  
  # Add custom HTML control for coordinates display
  htmlwidgets::onRender("
    function(el, x) {
      var map = this;
      
      // Create coordinate display div
      var info = L.control({position: 'bottomleft'});
      
      info.onAdd = function (map) {
        this._div = L.DomUtil.create('div', 'coordinate-info');
        this._div.style.padding = '6px 10px';
        this._div.style.background = 'rgba(255,255,255,0.9)';
        this._div.style.boxShadow = '0 0 15px rgba(0,0,0,0.2)';
        this._div.style.borderRadius = '5px';
        this._div.style.fontSize = '12px';
        this._div.style.fontFamily = 'monospace';
        this.update();
        return this._div;
      };
      
      info.update = function (lat, lng) {
        this._div.innerHTML = '<b>Coordinates:</b><br/>' +
          (lat && lng ? 
            'Lat: ' + lat.toFixed(5) + '°<br/>Lng: ' + lng.toFixed(5) + '°' : 
            'Hover over map');
      };
      
      info.addTo(map);
      
      // Update on mouse move
      map.on('mousemove', function(e) {
        info.update(e.latlng.lat, e.latlng.lng);
      });
      
      map.on('mouseout', function(e) {
        info.update();
      });

      // --- ENHANCED LAYER CONTROL ---
      setTimeout(function() {
        var layerControl = document.querySelector('.leaflet-control-layers-overlays');
        if (!layerControl) return;

        // 1. Add Toggle All Buttons
        var toggleBox = document.createElement('div');
        toggleBox.style.padding = '8px 12px';
        toggleBox.style.borderBottom = '1px solid #eee';
        toggleBox.style.marginBottom = '8px';
        toggleBox.style.backgroundColor = '#f8f9fa';
        toggleBox.style.textAlign = 'center';
        toggleBox.innerHTML = 
          '<button id=\"show-all-btn\" style=\"background:#fff; border:1px solid #ccc; border-radius:3px; font-size:11px; cursor:pointer; margin-right:5px; padding:2px 8px; color:#2c3e50;\">Show All</button>' +
          '<button id=\"hide-all-btn\" style=\"background:#fff; border:1px solid #ccc; border-radius:3px; font-size:11px; cursor:pointer; padding:2px 8px; color:#2c3e50;\">Hide All</button>';
        
        layerControl.parentNode.insertBefore(toggleBox, layerControl);

        var toggleLayers = function(check) {
          var inputs = layerControl.getElementsByTagName('input');
          for (var i = 0; i < inputs.length; i++) {
            if (inputs[i].checked !== check) inputs[i].click();
          }
        };

        document.getElementById('show-all-btn').onclick = function(e) { e.preventDefault(); e.stopPropagation(); toggleLayers(true); };
        document.getElementById('hide-all-btn').onclick = function(e) { e.preventDefault(); e.stopPropagation(); toggleLayers(false); };

        // 2. Insert Section Headers
        function addHeader(text, icon, beforeLayerName) {
          var labels = layerControl.getElementsByTagName('label');
          for (var i = 0; i < labels.length; i++) {
            if (labels[i].innerText.trim().includes(beforeLayerName)) {
              var header = document.createElement('div');
              header.innerHTML = '<div style=\"font-weight:bold; color:#2c3e50; margin:10px 0 5px 0; padding-bottom:3px; border-bottom:2px solid #3498db; font-size:12px; text-transform:uppercase;\">' + icon + ' ' + text + '</div>';
              labels[i].parentNode.insertBefore(header, labels[i]);
              break;
            }
          }
        }

        addHeader('Inventory', '📁', 'Zones');
        addHeader('Conservation', '🛡️', 'CFMAs');
        addHeader('Industrial', '🏭', 'Oil & Gas Blocks');
        addHeader('Fisheries', '🎣', 'Landing Sites');
        addHeader('Habitats', '🌿', 'Coral/Algae');
        addHeader('Analysis', '📊', 'Scenario');
      }, 100);

      // 3. Sync Legends with Layer Visibility
      map.on('overlayadd overlayremove', function() {
        var layers = map._layers;
        var showBenthicLegend = false;
        var benthicClasses = ['Coral/Algae', 'Seagrass', 'Microalgal Mats', 'Rock', 'Rubble', 'Sand'];
        
        for (var id in layers) {
          if (layers[id].options && layers[id].options.group) {
            if (benthicClasses.includes(layers[id].options.group)) {
              showBenthicLegend = true;
              break;
            }
          }
        }
        
        var bLegend = document.querySelector('.benthic-legend');
        if (bLegend) {
          bLegend.style.display = showBenthicLegend ? 'block' : 'none';
        }
      });
      
      // Trigger once on load
      setTimeout(function() {
        map.fire('overlayremove');
      }, 200);
    }
  ")

map

map |> htmlwidgets::saveWidget("index.html", selfcontained = T)

```

```{r}

```
